### 2018-06-20 parts/00_acknowledgements.tex
Added official LaTeX thesis class file, fixes issue #1

\begin{acknowledgments}

    \noindent
    We would like to thank Dr. Ma. Mercedes Rodrigo, Dr. Ma. Regina Estuar, and Dr. Proceso Fernandez, Jr. for initiating the creation of this thesis template for use of undergraduate thesis groups for years to come. We also extend our gratitude to Ms. Jessica Sugay for help invaluable help in automating many of the formatting of this template, especially in the creation of the custom styles, table of contents, list of figures, list of tables, and the bibliography.

    This section, as the name suggests, is the place in your paper where you may acknowledge individuals or groups who, with their help or guidance, made your study feasible and ultimately a reality. Some of these people include your adviser, volunteers for your study, other professors who may have contributed to your study, and if applicable, any group or organization who provided support in any way for your research.

\end{acknowledgments}