### 2018-06-20 parts/01_introduction_00.tex

\chapter{INTRODUCTION}

In this section of the paper, you want to provide the general background and motivation of your study. This can be done defining some key definitions or ideas that govern your study, presenting your context, and narrowing it down with parameters that shall concretize your study. 
This chapter is divided into four parts: the Context of the Study, the Research Objective/s, the Research Questions, the Scope and Limitations, and finally, the Significance of the Study. Each will be described in greater detail in their respective sections. Appendix A summarizes major points for each of these sections and gives sample guide questions for your reference.

\input{parts/01_introduction_01contextofstudy}
\input{parts/01_introduction_02researchquestions.tex}
\input{parts/01_introduction_03objectives.tex}
\input{parts/01_introduction_05scopeandlimitations}
\input{parts/01_introduction_04significanceofstudy}