### 2018-06-20 parts/05_conclusions.tex

\chapter{CONCLUSIONS}
Upon presenting your results, the conclusion is where you will now tie up these results with the original intent of the study, as indicated by the research questions given in the Introduction. It is in this section where you will also discuss any difficulties or issues encountered during the study, as well as your recommended method for addressing these problems.

The general way to organize your conclusion is to present each research sub-question as a subsection, and thoroughly answer each of them by interpreting your results with respect to the question. With these answered, you may then tie up all of your findings in each subsection to answer your main research question, providing any needed additional information or explanation. Last on the list would be your unsolved issues and difficulties, presenting them as avenues to motivate continued work on your chosen topic.