### 2018-06-20 thesis.tex

\documentclass[oneside,12pt]{discsugthesis}

\begin{document}
\ThesisAuthor{John H. Smith}
\ThesisTitle{THIS IS A THESIS TITLE}
\ThesisDegree{Computer Science}
\ThesisDefenseYear{2019}
\DefenseDate{}
\DepartmentHead{The Head}
\SchoolHead{The Head 2}
\FirstThesisAdviser{Ayun}
\FirstPanelMember{}
\SecondPanelMember{}
\ThirdPanelMember{}

\FrontMatter
\ThesisTitlePage
\input{parts/abstract.tex}
\tableofcontents
\listoffigures

\MainMatter
\input{parts/chapter1.tex}

\bibliography{thesis_refs}
\bibliographystyle{acm}
\nocite{*}
\end{document}