### 2018-06-25 parts/01_introduction_02researchquestions.tex

\section{Research Questions}

In this study, we ask the following main research question: What is the best method for image data to be encrypted and manipulated for practical applications?

To answer this question, we will consider the following sub-questions in our research:
\begin{enumerate}
	\item Which homomorphic image encryption algorithms are the most applicable for image processing operations?
	\item Which homomorphic image encryption algorithms are the most efficient in terms of time when applying image manipulation operations?
	\item Which homomorphic image encryption algorithms are the most secure under differential and entropy attacks?
\end{enumerate}