### 2018-06-25 parts/01_introduction_04significanceofstudy.tex

\section{Significance of the Study}

This study aims to directly address one of the current problems in the research of homomorphic cryptosystems: the practicality of homomorphic encryption \cite{sen_homomorphic_2013}. The literature shows existing homomorphic cryptosystems and how they support primitive operations (usually addition and multiplicaiton) on encrypted data, however, for homomorphic cryptosystems to be usable for practical image processing applications, support for more complicated operations such as image convolution must be demonstrated as well. While implementations of homomorphic encryption algorithms exist \cite{ziad_cryptoimg:_2016, garay_algorithms_2014}, our study targets algorithms for which statistical tests were not performed for image processing operations.

By developing a library allowing for the use and comparison of various homomorphic encryption algorithms, we wish to contribute to existing models for the secure transmission and modification of image data. Examples include applications in cloud storage \cite{potey_homomorphic_2016} and video surveillance \cite{upmanyu_efficient_2009}. This allows service providers to perform image manipulation for clients without compromising data privacy.