### 2018-07-14 parts/01_introduction_02researchquestions.tex

\section{Research Questions}

By implementing homomorphic cryptosystems and creating an OpenCV library so they can be easily used, we aim to address our main research question: What is the best method for image data to be encrypted and manipulated for practical applications?

More specifically, our research will address the following sub-research questions:
\begin{enumerate}
	\item Which homomorphic cryptosystems admit more image processing operations?
	\item Which homomorphic cryptosystems are the most time-efficient in applying image processing operations?
	\item Which homomorphic cryptosystems are the most secure under differential and entropy attacks?
\end{enumerate}