### 2018-07-14 parts/01_introduction_03objectives.tex

\section{Research Objectives}

A paper published in 2009 by Ziad, et. al. presents \textit{CryptoImg}, a library for the  Open Source Computer Vision Library (OpenCV) \cite{bradski_opencv_2000} which implements various homomorphic encryption and image processing routines using the Paillier homomorphic cryptosystem \cite{ziad_cryptoimg:_2016}. While \textit{CryptoImg} demonstrates that various image manipulation operations are possible in a homomorphic cryptosystem, we believe this study can be extended by considering other known homomorphic encryption schemes, such as those found in \cite{li_elliptic_2012, hutchison_fully_2010}. Our study aims to provide an objective comparison of various homomorphic cryptosystems currently in the literature, as they would be used in encrypted image manipulation.

Our main research objective is to assess the practicality of various homomorphic cryptosystems in image encryption. Homomorphic cryptosystems are known to be slower in terms of encryption, operation, and decryption time \cite{sen_homomorphic_2013}. Furthermore, the operations on encrypted data supported by homomorphic cryptosystems are limited, and some schemes require preprocessing to be performed on images, limiting the range of operations which can be performed \cite{li_elliptic_2012}. Finally, as data security is the primary purpose of cryptosystems, we want to evaluate the security of homomorphic cryptosystems as image encryption schemes.

In summary, in this study, we wish to
\begin{enumerate}
    \item Compare the applicability of known homomorphic encryption schemes to common image processing operations, in other words, determine which schemes admit more image processing operations;
		\item Compare the security of common image processing operations under known homomorphic encryption schemes, using established statistical methods in image encryption;
    \item Compare the time efficiency of common image processing operations under known homomorphic encryption schemes;
    \item Create a library for OpenCV to allow for convenient use and comparison of these homomorphic encryption schemes.
\end{enumerate}