### 2018-09-25 parts/02_relatedliterature_03facialrecognition.tex
New section for facial recognition

\section{Facial Recognition}
More complex image processing operations are just made up of simpler image operations such as those mentioned earlier. 
A prominent application of image processing is facial recognition.