### 2018-09-30 parts/01_introduction_02researchquestions.tex
[WIP] Restructuring RRL for midterms
Currently restructuring the RRL to accomodate algorithms, facial recognition.

\section{Research Questions}

By implementing homomorphic cryptosystems and creating an OpenCV library so they can be easily used, we aim to address our main research question: What is the best method for image data to be encrypted and manipulated for practical applications?

More specifically, our research will address the following sub-questions:
\begin{enumerate}
	\item Which homomorphic cryptosystems admit more image processing operations? How can existing homomorphic cryptosystems be modified to admit more image processing operations?
	\item Which homomorphic cryptosystems are the most time-efficient in applying image processing operations?
	\item Which homomorphic cryptosystems best preserve image quality in applying image processing operations? 
\end{enumerate}