### 2018-09-30 parts/02_relatedliterature.tex

\chapter{REVIEW OF RELATED LITERATURE}

We begin with an overview of image processing, in which we define the image processing operations we are interested in
We first introduce concepts in cryptography and define homomorphic cryptosytems. along with some examples from the literature. Then, we present some concepts in , a suitable application of cryptography, specifically the common image operations used. Finally, we present related work that attempts to apply image processing operations with homomorphic cryptosystems, as well as their advantages and limitations.

\input{parts/02_relatedliterature_01commonimageprocessingoperations}
%facial recognition - theory
\input{parts/02_relatedliterature_02homomorphiccryptosystems}
\input{parts/02_relatedliterature_03partiallyhomomorphic}
\input{parts/02_relatedliterature_04fullyhomomorphic}
%facial recognition - implementation
\input{parts/02_relatedliterature_05previousimplementations}

\section{Summary}
We have seen how various image processing operations make use of a series of additions and multiplications. Because of this, it is possible to use a homomorphic cryptosystem in order to apply image operations directly on encrypted images, as demonstrated by the \textit{CryptoImg} library.

In particular, we aim to address the shortcomings of \textit{CryptoImg} by developing a library to test various homomorphic encryption schemes with the help of the \textit{HElib} library to be able to implement fully homomorphic encryption schemes efficiently. This helps us determine the practicality of using homomorphic cryptosystems in image processing, especially when dealing with more complex image operations.