### 2018-10-05 parts/02_relatedliterature.tex
Finalized new ordering of RRL
Restructured RRL to accomodate facial recognition

\chapter{REVIEW OF RELATED LITERATURE}

We will now provide an overview of the \textit{CryptoImg} project, which our study is primarily based on. Then, we define the image processing operations we are interested in implementing in this study and
After establishing our desired image processing operations, introduce concepts in cryptography and define the four homomorphic cryptosystems we will be testing in the study.

\input{parts/02_relatedliterature_01cryptoimg}
\input{parts/02_relatedliterature_02commonimageprocessingoperations}
%facial recognition - theory
\input{parts/02_relatedliterature_03facialrecognition}
\input{parts/02_relatedliterature_04homomorphiccryptosystems}
\input{parts/02_relatedliterature_05partiallyhomomorphic}
\input{parts/02_relatedliterature_06fullyhomomorphic}
%facial recognition - implementation
\input{parts/02_relatedliterature_07previousimplementations}

\section{Summary}
We have seen how various image processing operations make use of a series of additions and multiplications. Because of this, it is possible to use a homomorphic cryptosystem in order to apply image operations directly on encrypted images, as demonstrated by the \textit{CryptoImg} library.

In particular, we aim to address the shortcomings of \textit{CryptoImg} by developing a library to test various homomorphic encryption schemes with the help of the \textit{HElib} library to be able to implement fully homomorphic encryption schemes efficiently. This helps us determine the practicality of using homomorphic cryptosystems in image processing, especially when dealing with more complex image operations.