### 2018-10-05 parts/02_relatedliterature_03facialrecognition.tex

\section{Facial Detection and Recognition}
%% TODO: Brian dump ur stuff here