### 2018-10-07 parts/02_relatedliterature.tex
RRL summary and fixes
- Commented out spatial filter section
- Updated summary to reflect new research
- Fixed some typesetting errors in D-P section

\chapter{REVIEW OF RELATED LITERATURE}

We will now provide an overview of the \textit{CryptoImg} project, which our study is primarily based on. Then, we define the image processing operations we are interested in implementing in this study and
After establishing our desired image processing operations, introduce concepts in cryptography and define the four homomorphic cryptosystems we will be testing in the study.

\input{parts/02_relatedliterature_01cryptoimg}
\input{parts/02_relatedliterature_02commonimageprocessingoperations}
%facial recognition - theory
\input{parts/02_relatedliterature_03facialrecognition}
\input{parts/02_relatedliterature_04homomorphiccryptosystems}
\input{parts/02_relatedliterature_05partiallyhomomorphic}
\input{parts/02_relatedliterature_06fullyhomomorphic}
%facial recognition - implementation
\input{parts/02_relatedliterature_07previousimplementations}

\section{Summary}
We have seen how privacy-preserving image operations have been implemented in the literature using partially homomorphic cryptosystems. \textit{CryptoImg} provided a suite of linear image processing operations using the Paillier cryptosystem, while Erkin, et al. have implemented a privacy-preserving version of the eigenfaces approach to facial recognition developed by Turk and Pentland.

We have presented as well several areas in which our study will expand on the research so far.
\begin{enumerate}
  \item We have shown two fully homomorphic cryptosystems, the Dasgupta-Pal cryptosystem and the BGV cryptosystem. While these cryptosystems are generally less efficient than partially homomorphic cryptosystems, they do not require additional protocols for exponentiation.
  \item Due to the general limitations of partially homomorphic systems, attempts at applying homomorphic cryptography to applications such as image processing and facial recognition have mainly been restricted to linear operations on ciphertext. We have shown that there are basic image processing operations which are non-linear, which we can implement.
  \item While privacy-preserving facial recognition has been accomplished by Erkin, et al., privacy-preserving facial detection is an area of research which has yet to explored.
\end{enumerate}