### 2019-01-09 paper/PCSC2019-sigconf.tex
Prepare thesis for PCSC

\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage{multicol}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs} % For formal tables

\setcopyright{rightsretained}

% DOI
\acmDOI{}

% ISBN
\acmISBN{}

%Conference
\acmConference[PCSC2019]{Philippine Computing Science Congress}{March 2019}{City of Manila, Philippines}
\acmYear{2019}
\copyrightyear{2019}

\acmArticle{}
\acmPrice{}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
\editor{}
\editor{}
\editor{}

\begin{document}
\title{Implementation and Analysis of Homomorphic Facial Image Encryption and Manipulation}

\author{Me 1}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin}
  \state{Ohio}
  \postcode{43017-6221}
}
\email{webmaster@marysville-ohio.com}

\author{Me 2}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

% The default list of authors is too long for headers.
\renewcommand{\shorttitle}{Implementation and Analysis}
% \renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
This paper provides a sample of a \LaTeX\ document which conforms,
somewhat loosely, to the formatting guidelines for
ACM SIG Proceedings.\footnote{This is an abstract footnote}
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
  <ccs2012>
  <concept>
  <concept_id>10002978.10002979.10002981.10011745</concept_id>
  <concept_desc>Security and privacy~Public key encryption</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10002978.10003018.10003020</concept_id>
  <concept_desc>Security and privacy~Management and querying of encrypted data</concept_desc>
  <concept_significance>300</concept_significance>
  </concept>
  <concept>
  <concept_id>10010147.10010371.10010382.10010383</concept_id>
  <concept_desc>Computing methodologies~Image processing</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10002950.10003714.10003740</concept_id>
  <concept_desc>Mathematics of computing~Quadrature</concept_desc>
  <concept_significance>100</concept_significance>
  </concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Public key encryption}
\ccsdesc[300]{Security and privacy~Management and querying of encrypted data}
\ccsdesc[500]{Computing methodologies~Image processing}
\ccsdesc[100]{Mathematics of computing~Quadrature}

\keywords{ACM proceedings, \LaTeX, text tagging}

\maketitle

\input{paper-conf}

\appendix
\input{appendix}

\begin{acks}
  The authors would like to thank Dr. Yuhua Li for providing the
  MATLAB code of the \textit{BEPS} method.

  The authors would also like to thank the anonymous referees for
  their valuable comments and helpful suggestions. The work is
  supported by the \grantsponsor{GS501100001809}{National Natural
    Science Foundation of
    China}{http://dx.doi.org/10.13039/501100001809} under Grant
  No.:~\grantnum{GS501100001809}{61273304}
  and~\grantnum[http://www.nnsf.cn/youngscientists]{GS501100001809}{Young
    Scientists' Support Program}.

\end{acks}

\bibliographystyle{ACM-Reference-Format}
\bibliography{thesis_refs}

\end{document}