### 2019-01-16 pcsc-paper/parts/06_conclusionreccomendations.tex

\section{Conclusion [WIP]}
\section{Recommendations [WIP]}