### 2019-01-18 pcsc-paper/parts/06_conclusionreccomendations.tex

\section{Conclusion}
We have shown new closed-form approximations to the logarithm and arctangent functions and showed how they can be implemented together with the Paillier cryptosystem to allow for privacy-preserving function calculation. We have compared our approach to other closed-form approximations which require a similar number of secure arithmetic operations to evaluate, and have shown our new approximation procedures can yield more accurate results over larger ranges of inputs.

\section{Recommendations}
Further research can be done regarding implementation of other transcendental functions, either by using known infinite series, or by using identities with the quadrature-based approximations we have presented in this paper. Furthermore, testing with other types of homomorphic cryptosystems can be done, to potentially reduce noise encountered in secure floating-point operations.