### 2019-01-18 pcsc-paper/parts/00_abstract.tex
Paper almost finished

This paper presents methods to securely compute transcendental functions such as the natural logarithm and the inverse tangent function within a homomorphic cryptosystem.
First, an extension to the Paillier cryptosystem that allows operations on floating-point numbers is presented. 
Second, protocols using a two-party scheme are described for secure multiplication, division, and exponentiation, in order for such operations to be supported by a partially homomorphic cryptosystem such as the Paillier cryptosystem, where these operations are theoretically infeasible to do under a single-party system.
Finally, new closed-form approximations to the logarithm and the inverse tangent functions are derived using the Gauss--Legendre quadrature, which yield more accurate results than other closed-form approximations that involve a similar number of secure operations.