### 2019-01-18 pcsc-paper/parts/05_testsforaccuracy.tex

\section{Experiments on Accuracy}
We implemented privacy-preserving floating-point arithmetic using the system described in Section~\ref{sec:fp_operations}, through the \texttt{python-paillier} library (\url{https://github.com/n1analytics/python-paillier}) running on Python~3.6.5.

\subsection{Logarithm Approximation}
We then computed $\log(1+x)$ for various encrypted inputs using both the original quadrature formula by Khattri (Equation~\ref{eq:standard_logarithm_quadrature}) and our scaled approximation (Equation~\ref{eq:optimal_log_approximation}). These were compared with the values returned by the built-in logarithm function in Python. Several data points are shown in Table~\ref{tab:log_approximation}.

We note that the privacy-preserving floating-point arithmetic resulted in some degree of floating-point error, as the observed floating point error in Table~\ref{tab:log_approximation} is greater than the raw absolute error observed in Figure~\ref{fig:log_error_comparison}. However, it can still be seen that the error of the scaled approximation is less than that of the original approximation for large inputs.
\begin{table}[h]
	\caption{Comparison of logarithm approximation errors under Paillier}
	\label{tab:log_approximation}
	\begin{tabular}{lcc}
		\toprule
		Input & Original quadrature & Scaled quadrature\\
		\midrule
		$\log 1$ & $0$ & $2.053729 \times 10^{-2}$\\
		$\log 10$ & $2.833179 \times 10^{-2} $ & $2.280124 \times 10^{-1}$\\
		$\log 50$ & $2.170779 \times 10^{-2}$ & $2.437148 \times 10^{-2}$\\
		$\log 100$ & $3.524916 \times 10^{-1}$ & $2.122159 \times 10^{-2}$\\
		$\log 200$ & $8.859869 \times 10^{-1}$ & $1.441611 \times 10^{-1}$\\
		$\log 256$ & $1.098463 \times 10^{0}$ & $1.993136 \times 10^{-1}$\\
	\bottomrule
\end{tabular}
\end{table}

\subsection{Inverse Tangent Approximation}
We also computed $\arctan x$ using for various encrypted inputs using both the first five terms of Euler's series for the inverse tangent function (Equation~\ref{eq:arctan_euler_partial}) and our quadrature approximation (Equation~\ref{eq:arctan_quadrature}). These were compared with the values returned by the built-in inverse tangent function in Python. We show several data points in Table~\ref{tab:arctan_approximation}. Similar to the logarithm approximations, we note that there is a degree of floating-point error compared to that predicted by Figure~\ref{fig:arctan_error}, but the quadrature approach obtains higher accuracy despite requiring a similar number of floating-point operations to execute.

\begin{table}[h]
	\caption{Comparison of inverse tangent approximation errors under Paillier}
	\label{tab:arctan_approximation}
	\begin{tabular}{lcc}
		\toprule
		Input & Partial sum of infinite series & Quadrature\\
		\midrule
		$\arctan -1$ & $6.564089 \times 10^{-1}$ & $4.629557 \times 10^{-1}$\\
		$\arctan -0.1$ & $4.523405 \times 10^{-3}$ & $1.451929 \times 10^{-3}$\\
		$\arctan -0.01$ & $4.665260 \times 10^{-6}$ & $1.472027 \times 10^{-6}$\\
		$\arctan 0$ & $0$ & $0$\\
		$\arctan 0.01$ & $4.665228 \times 10^{-6}$ & $1.472020 \times 10^{-6}$\\
		$\arctan 0.1$ & $4.523405 \times 10^{-3} $ & $1.451929 \times 10^{-3}$\\
		$\arctan 1$ & $6.564089 \times 10^{-1} $ & $4.629557 \times 10^{-1}$\\
	\bottomrule
\end{tabular}
\end{table}