### 2019-02-08 paper/icmssp2019-sigconf.tex
Started edits

% \documentclass[sigconf]{acmart}
\documentclass{sig-alternate-05-2015}
\usepackage{natbib}
% \usepackage[LGR,T1]{fontenc}‎
\usepackage[utf8]{inputenc}
\usepackage[hidelinks]{hyperref}
\usepackage{newtxtext}
\usepackage{zi4}
\usepackage{newtxmath}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xinttools}
\usepackage{array}
\usepackage{booktabs} % For formal tables
\usepackage{pgf, tikz}
\usepackage{pgfplots}
\usetikzlibrary{calc}

\setcopyright{acmlicensed}
\newtheorem{theorem}{Theorem}

% DOI
% \acmDOI{}

% ISBN
% \acmISBN{}

%Conference
% \acmConference[PCSC2019]{Philippine Computing Science Congress}{March 2019}{City of Manila, Philippines}
% \acmYear{2019}
\CopyrightYear{2019}

% \acmArticle{}
% \acmPrice{}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
% \editor{}
% \editor{}
% \editor{}

\begin{document}
\setcitestyle{numbers,square,comma,sort}
\title{Implementation and Analysis of Homomorphic Facial Image Encryption and Manipulation}

% \author{Aldrich Ellis C. Asuncion}
% \affiliation{%
%   \institution{Ateneo de Manila University}
%   \streetaddress{Katipunan Avenue, Loyola Heights}
%   \city{Quezon City}
%   \country{Philippines}
%   \postcode{1108}
% }
% \email{aldrich.asuncion@obf.ateneo.edu}

% \author{Brian Christopher T. Guadalupe}
% \affiliation{%
%   \institution{Ateneo de Manila University}
%   \streetaddress{Katipunan Avenue, Loyola Heights}
%   \city{Quezon City}
%   \country{Philippines}
%   \postcode{1108}
% }
% \email{brian.guadalupe@obf.ateneo.edu}

\numberofauthors{2}
\author{
\alignauthor Aldrich Ellis C. Asuncion\\
\affaddr{Ateneo de Manila University}\\
\affaddr{Katipunan Avenue, Loyola Heights}\\
\affaddr{Quezon City, Philippines}\\
\email{aldrich.asuncion@obf.ateneo.edu}
\alignauthor Brian Christopher T. Guadalupe\\
\affaddr{Ateneo de Manila University}\\
\affaddr{Katipunan Avenue, Loyola Heights}\\
\affaddr{Quezon City, Philippines}\\
\email{brian.guadalupe@obf.ateneo.edu}
}

\maketitle

\begin{abstract}
Homomorphic cryptography allows for encrypted data to be modified and operated on without requiring decryption, although current homomorphic cryptosystems are either limited in their permitted operations or are significantly more time-intensive than standard non-homomorphic cryptosystems. Regardless, homomorphic cryptography has been targeted for use in secure image processing and facial recognition, due to their ability to maintain data privacy. In this paper, we compared the viability of the Paillier, Damg{\aa}rd--Geisler--Kr{\o}igaard (DGK), Dasgupta--Pal, and Brakerski--Gentry--Vaikuntanathan (BGV) cryptosystems for facial image processing applications, by implementing a software library equipped with these cryptosystems, and comparing their time efficiency and accuracy. Furthermore, as an extension to previous research, we attempt to support non-linear image processing operations. Preliminary results have shown that the Paillier and Dasgupta--Pal cryptosystems are comparable in accuracy and may be used to approximate a logarithm transformation, but the Dasgupta--Pal cryptosystem is significantly slower.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below.
%
\begin{CCSXML}
  <ccs2012>
  <concept>
  <concept_id>10002978.10002979.10002981.10011745</concept_id>
  <concept_desc>Security and privacy~Public key encryption</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10002978.10003018.10003020</concept_id>
  <concept_desc>Security and privacy~Management and querying of encrypted data</concept_desc>
  <concept_significance>300</concept_significance>
  </concept>
  <concept>
  <concept_id>10002978.10003029.10011150</concept_id>
  <concept_desc>Security and privacy~Privacy protections</concept_desc>
  <concept_significance>300</concept_significance>
  </concept>
  <concept>
  <concept_id>10010147.10010371.10010382.10010383</concept_id>
  <concept_desc>Computing methodologies~Image processing</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Security and privacy~Public key encryption}
\ccsdesc[300]{Security and privacy~Management and querying of encrypted data}
\ccsdesc[300]{Security and privacy~Privacy protections}
\ccsdesc[500]{Computing methodologies~Image processing}

\printccsdesc
\keywords{cryptography; homomorphic encryption; data privacy; image processing; intensity transformation; facial recognition}

% \maketitle

\input{paper-conf}

% \begin{acks}
\section{Acknowledgments}
% We would like to thank our adviser Dr. William Emmanuel S. Yu for his guidance and constant feedback, and 
We would like to thank our panelists Dr. Andrei D. Coronel and Dr. Mari-Jo P. Ruiz for lending their expertise and for giving their invaluable comments and insights.
Due acknowlegement is also given to the University of Essex for making the \texttt{faces94} dataset publicly available for research purposes.
% \end{acks}

\bibliographystyle{ACM-Reference-Format}
% \citestyle{acmnumeric}
% \bibliographystyle{acm}
\bibliography{thesis_refs}
\balancecolumns

% \appendix
% \input{appendix}
% \balancecolumns

\end{document}