### 2019-02-28 thesis.tex

\documentclass[oneside,12pt]{discsthesis}
\usepackage{graphicx}
\usepackage{amsthm}
%\usepackage{multirow}
%\usepackage{fancyvrb}
\usepackage{tabularx}
\usepackage[titletoc]{appendix}
\usepackage{subcaption}
%%\usepackage{rotating}
%%\usepackage{pdfpages}
%\usepackage{lscape}
%\usepackage{adjustbox}
%\usepackage{longtable}
\usepackage{url}
\usepackage{verbatim}
\usepackage{pgf, tikz}
\usepackage{pgfplots}

% commands
\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newtheorem{theorem}{Theorem}[chapter]

\title{FinalManuscript_DISCS_UG}
\author{acalab.imagegroup }
\date{March 1, 2018}

\begin{document}

\ThesisAuthor{Aldrich Ellis C. Asuncion \\ Brian Christopher T. Guadalupe}
\ThesisTitle{IMPLEMENTATION AND ANALYSIS OF HOMOMORPHIC FACIAL IMAGE ENCRYPTION AND MANIPULATION}
\ThesisArea{Computer Science}
\ThesisDefenseYear{2019}
\DefenseDate{March 1, 2018}

\DepartmentHead{ANDREI D. CORONEL, Ph.D.}
\SchoolHead{EVANGELINE P. BAUTISTA, Ph.D.}
\ThesisAdviser{MARLENE M. DE LEON, Ph.D.}
\FirstPanelMember{MA. MERCEDES T. RODRIGO, Ph.D.}
\SecondPanelMember{JOHN PAUL C. VERGARA, Ph.D.}
\ThirdPanelMember{MA. REGINA JUSTINA E. ESTUAR, Ph.D.}

\ThesisStyle{MS}{Draft}{-10pt}{15pt}

%% Choices for the 1st argument: MS, PhD
%% Choices for the 2nd argument: Final, FinalWithCorner, Draft

%\FrontMatter
% \input{parts/00_abstract}
% \input{parts/00_acknowledgements}
\tableofcontents
\listoffigures
% \listoftables

\MainMatter
\input{parts/01_introduction_00}
\input{parts/02_relatedliterature}
\input{parts/03_methodology}
% \input{parts/04_results}
%\input{parts/04_results_pub}
%\input{parts/04_results_unpub}
% \input{parts/05_conclusions}
%\input{parts/06_recommendations}

% BIBLIOGRAPHY
%\bibliographystyle{abbrv}
%\bibliography{sigproc}
%\input{parts/09_references}
%\BackMatter
\bibliographystyle{acm}
\bibliography{thesis_refs}

\addappheadtotoc
\begin{appendices}
% \input{parts/08_appendices_a_cryptosystems}
\input{parts/08_appendices_b_numericalapproximations_updated}
\end{appendices}

\end{document}