### 2019-04-29 parts/01_introduction_02researchquestions.tex

\section{Research Questions}

By creating a sofware library which implements various homomorphic cryptosystems, we aim to address our main research question: What is the best method for image data to be encrypted and manipulated for practical applications?

Specifically, the study addresses the following research sub-questions:
\begin{enumerate}
	\item Which homomorphic cryptosystems admit more image processing operations? How can existing homomorphic cryptosystems be modified to admit more image processing operations?
	\item Which homomorphic cryptosystems are the most time-efficient and the most accurate in applying image processing operations?
	\item Which homomorphic cryptosystems best preserve image quality in applying image processing operations?
\end{enumerate}