### 2019-04-30 parts/01_introduction_01contextofstudy.tex

\section{Context of the Study}

Digital data privacy and security is a growing concern in various fields, such as cloud computing~\cite{potey_homomorphic_2016}, health information systems~\cite{kester_cryptographic_2015}, and video surveillance~\cite{upmanyu_efficient_2009}. Various cryptosystems have been developed which allow for the secure transmission of data using encryption and decryption schemes~\cite{bauer_cryptosystem_2005}.

One particular application of cryptography is image encryption. While cryptosystems used to encrypt general digital data, such as the Advanced Encryption Standard (AES) and elliptic curve cryptography, can also be used to encrypt images~\cite{jain_image_2016, singh_image_2015} other algorithms created specifically for image encryption have also been developed~\cite{murugan_survey_2018}.

 Using a homomorphic cryptosystem, operations can be performed on encrypted data while keeping both the operands and resulting values secure. The result of the operations can then be recovered when the data is decrypted. In this manner, data is manipulated without compromising data privacy~\cite{fontaine_survey_2007, sen_homomorphic_2013}. There is additional interest in being able to perform image manipulation operations on encrypted images such as image adjustment, filtering, and morphological/feature extraction operations~\cite{ziad_cryptoimg:_2016, gonzalez_digital_2008}. More advanced operations have also been implemented using homomorphic cryptosystems, such as facial recognition \cite{hutchison_privacy-preserving_2009} and neural networks \cite{hesamifard_cryptodl:_2017}. While secure linear image operations have been implemented \cite{ziad_cryptoimg:_2016}, preprocessing which requires non-linear operations, such as certain methods of illumination processing \cite{oravec_illumination_2010}, has yet to be performed in the encrypted domain. This study in particular focuses on facial recognition.